### 2019-04-23 report/chapter03.Rnw
Meta analysis finish

<<echo=FALSE>>=
library(knitr)
opts_chunk$set(fig.path='figure/ch02_fig',
               echo=TRUE, message=FALSE,
               fig.width=8.1, fig.height=3.5,
               out.width='\\textwidth-3cm',
               message=FALSE, fig.align='center',
               background="gray98", tidy=FALSE, #tidy.opts=list(width.cutoff=60),
               cache=TRUE
)
options(width=74)
@

@@

\section{Meta-analysis}
It is possible to analyse all studies with one or more replica by meta-analysis. There are different meta-analyses that can be applied. First, it is investigated how well the significance of a effect size estimate of a study replicates in significance of the pooled effect size estimate of a meta-analysis. First, the amount of significant study effect estimates is shown over time. The fraction of significant effect size estimates on the $p$-value threshold of $0.05$ is only shown for times where a reasonbly large number of effect estimates is available in order to reduce random fluctuation ($n > 800$). Shown are p-values for significance tests where the outcome measure of the study was either a risk ratio, odds ratio, mean difference and standardized mean difference. Additionally, only study years between 1970 and 2017 have been included and the sample size for each arm of the study had to $> 12$.

\begin{figure}
<<>>=
data.ext %>% filter(!is.na(sig.type)) %>%  ggplot(aes(x = study.year, fill = factor(sig.type), stat(count))) + geom_density(na.rm = T, position = "fill") + 
  labs(fill = "Significance") + scale_fill_discrete(labels= c("Yes", "No"))+ scale_x_continuous(limits = c(1970, 2017))
@
\caption{Mean of the absolute value of the normalized effect size plotted against the total sample size.}
\label{study.significance.overtime}
\end{figure}

@@

\begin{figure}
<<echo=FALSE>>=
data %>% filter(total1 + total2 > 10 & total1 + total2 < 200) %>% 
  mutate(sample.size = total1 + total2, scaled.effect = abs(scale(effect, center = T, scale = T))) %>%
  group_by(sample.size) %>% 
  summarize(median.effect = median(scaled.effect, na.rm = T)) %>% 
  ggplot(aes(x = sample.size, y = median.effect)) + geom_point() + theme_bw() + xlab("sample size") + ylab("median absolute normalized effect size")
@
\caption{Mean of the absolute value of the normalized effect size plotted against the total sample size.}
\label{effect.samplesize}
\end{figure}

@@

\begin{figure}
<<echo=FALSE>>=
data %>% filter(total1 + total2 > 10 & total1 + total2 < 100) %>% 
  filter(outcome.measure.new == "Risk Ratio" | outcome.measure.new == "Odds Ratio" | outcome.measure.new == "Mean Difference" |
           outcome.measure.new == "Std. Mean Difference") %>% 
  group_by(outcome.measure.new) %>% 
  mutate(sample.size = total1 + total2, scaled.effect = abs(scale(effect, center = T, scale = T))) %>%
  group_by(sample.size, outcome.measure.new) %>% 
  summarize(median.effect = median(scaled.effect, na.rm = T)) %>% 
  ggplot(aes(x = sample.size, y = median.effect, group = outcome.measure.new)) + geom_point() + facet_wrap(~outcome.measure) + 
  theme_bw() + xlab("sample size") + ylab("median absolute normalized effect size")
@
\caption{Mean of the absolute value of the normalized effect size plotted against the total sample size, separated for outcome measures.}
\label{effect.samplesize.separated}
\end{figure}

A second way to analyse meta-analyses is a cumulative meta-analysis that should reveal shifts in treatment effect sizes over time. This can again be done for the entire dataset, i.e. for all effect estimates and their ordering in time. It is important to scale the effect estimates here with respect to the estimates of the replication studies that are about the same subject, have the same outcome measure, etc. Also, only effect estimates that can be compared to other estimates before are included, i.e. only study results with one or more replica are included. Time needs to be scaled and normalized in order to compare multiple time-trends in effect size to each other and gain insights in the overall trend. This is done in Figure \ref{effects.over.time}. In blue, the fit of a linear regression is depicted with grey 95 \% confidence intervals.

\begin{figure}
<<echo=FALSE>>=
data %>% filter(total1 + total2 > 10 & total1 + total2 < 100) %>% filter(!is.na(study.year)) %>% 
  filter(outcome.measure.new == "Risk Ratio" | outcome.measure.new == "Odds Ratio" | outcome.measure.new == "Mean Difference" |
           outcome.measure.new == "Std. Mean Difference") %>% 
  group_by(outcome.measure.new, comparison.nr, outcome.nr, subgroup.nr) %>% count() %>% filter(n > 1) %>% 
  mutate(scaled.effect = scale(effect), scaled.time = scale(study.year)) %>% 
  ggplot(aes(x = scaled.time, y = scaled.effect)) + geom_point(size = .5, alpha = 0.2)  + 
  geom_smooth(method = lm) + theme_bw() + 
  labs(title = "Scaled and centered RR over time", subtitle = "Initial finding > 1") + ylab("Risk Ratio")  + ylab("Time")
@
\caption{Mean of the absolute value of the normalized effect size plotted against the total sample size, separated for outcome measures.}
\label{effect.samplesize.separated}
\end{figure}

% It is also possible to look how the amount of meta-analysis with significant pooled effect size estimates changes over time, e.g. mean publication year (of studies included in the meta-analysis). 0.05 is used here as the threshold for $p$-value significance and subsequent null-hypothesis rejection. To constrict random fluctuation in the proportion, only mean publication years with a certain amount of meta-analyses are shown ($n < 180$). The $p$-value of the fixed effects meta analysis model is used to decide about rejection of the null-hypothesis. As Figure \ref{treatment.effect.significance.overtime} shows, the proportion of meta-analyses with significant treatment effect estimates stays roughly constant over time. 
% 
% \begin{figure}
% <<echo=FALSE>>=
% meta %>% filter(mean.publication.year < 2013 & mean.publication.year > 1990) %>% 
%   ggplot(aes(x = mean.publication.year, fill = factor(sig.fixef), stat(count))) + geom_density(na.rm = T, position = "fill") + 
%   labs(fill = "Significance") + scale_fill_discrete(labels= c("Yes", "No"))
% @
% \caption{Proportion of significant treatment effect estimates over time (mean publication year).}
% \label{effect.samplesize}
% \end{figure}
% 
% A different question is how statistical power of studies changed over time. As an approximation, the total sample size of a study can be used. Figure \ref{median.samplesize.overtime} shows the median sample size of all studies of a given year. Again, to reduce the effects of distortion due to random fluctuation, only the medians for the years 1980 to 2017 are used ($n$ > 50,000).
% 
% \begin{figure}
% <<echo=FALSE>>=
% data %>% distinct(study.name, .keep_all = T) %>%  filter(study.year < 2018 & study.year > 1980) %>% group_by(study.year) %>% 
%   summarize(samplesize = median(total1 + total2, na.rm = T)) %>% 
%   ggplot(aes(y = samplesize, x = study.year)) + geom_line()
% @
% \caption{Median sample size over time (years).}
% \label{median.samplesize.overtime}
% \end{figure}

\subsection{Meta-analyses}

Meta-analyses can be applied on the dataset given that there are studies that have one or more replica (i.e. same comparison type, outcome type and subgroup type).

@@

p.change.ranef.bin <- meta.bin %>% ungroup() %>% 
  select(sig.change.ranef.reg, sig.change.ranef.copas, sig.change.ranef.trimfill) %>% 
  gather(key = "correction.method", value = "sig.change") %>%  
  mutate(sig.change = factor(sig.change)) %>% 
  ggplot(aes(x = correction.method, fill = sig.change)) + geom_bar() + coord_flip() + ggtitle("Binary outcomes")+ 
  theme_bw() + xlab(label = NULL) + scale_fill_discrete(labels= c("unchanged.nonsig", "change.to.sig", "unchanged.sig", "change.to.nonsig")) +
  annotate("text", x = change.ranef.bin[change.ranef.bin$sig.change == 3,]$correction.method, y = 300, 
         label = paste(change.ranef.bin[change.ranef.bin$sig.change == 3,]$n, "changed to nonsig."), 
         color = "white")

@@

p.change.ranef.cont <- meta.cont %>% ungroup() %>% 
  select(sig.change.ranef.reg, sig.change.ranef.copas, sig.change.ranef.trimfill) %>% 
  gather(key = "correction.method", value = "sig.change") %>%  
  mutate(sig.change = factor(sig.change)) %>% 
  ggplot(aes(x = correction.method, fill = sig.change)) + geom_bar() + coord_flip() + ggtitle("Continuous outcomes") +
  theme_bw() + xlab(label = NULL) + scale_fill_discrete(labels= c("unchanged.nonsig", "change.to.sig", "unchanged.sig", "change.to.nonsig")) +
  annotate("text", x = change.ranef.cont[change.ranef.cont$sig.change == 3,]$correction.method, y = 200, 
         label = paste(change.ranef.cont[change.ranef.cont$sig.change == 3,]$n, "changed to nonsig."), 
         color = "white")

@@

p.ranef.change <- meta %>% ungroup() %>% 
  select(sig.change.ranef.reg, sig.change.ranef.copas, sig.change.ranef.trimfill) %>% 
  gather(key = "correction.method", value = "sig.change") %>%  
  mutate(sig.change = factor(sig.change)) %>% 
  ggplot(aes(x = correction.method, fill = sig.change)) + geom_bar() + coord_flip() + 
  theme_bw() + xlab(label = NULL) + scale_fill_discrete(labels= c("unchanged.nonsig", "change.to.sig", "unchanged.sig", "change.to.nonsig")) +
  annotate("text", x = change.ranef[change.ranef$sig.change == 3,]$correction.method, y = 500, 
         label = paste(change.ranef[change.ranef$sig.change == 3,]$n, "changed to nonsig."), 
         color = "white")

@@

p.change.fixef.bin <- meta.bin %>% ungroup() %>% 
  select(sig.change.fixef.reg, sig.change.fixef.copas, sig.change.fixef.trimfill) %>% 
  gather(key = "correction.method", value = "sig.change") %>%  
  mutate(sig.change = factor(sig.change)) %>% 
  ggplot(aes(x = correction.method, fill = sig.change)) + geom_bar() + coord_flip() + ggtitle("Binary outcomes")+  
  theme_bw() + xlab(label = NULL) + scale_fill_discrete(labels= c("unchanged.nonsig", "change.to.sig", "unchanged.sig", "change.to.nonsig")) +
  annotate("text", x = change.fixef.bin[change.fixef.bin$sig.change == 3,]$correction.method, y = 200, 
         label = paste(change.fixef.bin[change.fixef.bin$sig.change == 3,]$n, "changed to nonsig."), 
         color = "white")

@@

p.fixef.change <- meta %>% ungroup() %>% 
  select(sig.change.fixef.reg, sig.change.fixef.copas, sig.change.fixef.trimfill) %>% 
  gather(key = "correction.method", value = "sig.change") %>%  
  mutate(sig.change = factor(sig.change)) %>% 
  ggplot(aes(x = correction.method, fill = sig.change)) + geom_bar() + coord_flip() + 
  theme_bw() + xlab(label = NULL) + scale_fill_discrete(labels= c("unchanged.nonsig", "change.to.nonsig", "unchanged.sig", "change.to.sig")) +
  annotate("text", x = change.fixef[change.fixef$sig.change == 3,]$correction.method, y = 500, 
         label = paste(change.fixef[change.fixef$sig.change == 3,]$n, "changed to nonsig."), 
         color = "white")

@@

\subsubsection{Small Study Effect Tests}
There are tests that can be applied to find out if small study effects are present in the meta analysis. For the precise description, see the methods section. Application of the tests is only recommended if there are ten or more studies \citep{cochrane.handbook} that can be used, so all meta-analyses with less than ten studies have been excluded.

@@

\begin{figure}
<<echo=FALSE>>=
plot(p.cont)
@
\caption{Proportion where the null hypothesis of no small study effect is rejected based on the 5\% significance level for different tests (continuous outcomes).}
\label{bias.results.cont}
\end{figure}

\begin{figure}
<<echo=FALSE>>=
plot(p.bin)
@
\caption{Proportion where the null hypothesis of no small study effect is rejected based on the 5\% significance level for different tests (dichotomous outcomes).}
\label{bias.results.bin}
\end{figure}

@@

\begin{figure}
<<echo=FALSE>>=
grid.arrange(p3, p4, ncol = 2)
@
\caption{Proportion where the null hypothesis of no small study effect is rejected based on the 5\% significance level for different tests (continuous outcomes).}
\label{bias.results.cont.sep}
\end{figure}

\begin{figure}
<<echo=FALSE>>=
grid.arrange(p1, p2, ncol = 2)
@
\caption{Proportion where the null hypothesis of no small study effect is rejected based on the 5\% significance level for different tests (dichotomous outcomes).}
\label{bias.results.bin.sep}
\end{figure}

@@

\begin{figure}
<<echo=FALSE>>=
meta %>% filter(mean.publication.year < 2013 & mean.publication.year > 1990) %>% 
  ggplot(aes(x = mean.publication.year, fill = factor(thomson.test), stat(count))) + geom_density(na.rm = T, position = "fill") + 
  labs(fill = "Significance") + scale_fill_discrete(labels= c("Yes", "No"))
@
\caption{Proportion of test results where the null hypothesis of no small study effect is rejected over time (mean publication.year).}
\label{pub.bias.overtime}
\end{figure}

@@

<<results = 'asis', echo = FALSE, warnings = FALSE>>=
print(xtable(x = t(binary.tests.agreement), caption = "Proportion of tests that agree in rejection ar acceptance of the null hypothesis that there is no small study effect (dichotomous outcomes)", label = "agreement.bin",  align = "lll", digits = 2), include.rownames = T, size = "footnotesize")
@

<<results = 'asis', echo = FALSE, warnings = FALSE>>=
print(xtable(x = t(cont.tests.agreement), caption = "Proportion of tests that agree in rejection ar acceptance of the null hypothesis that there is no small study effect (continuous outcomes)", label = "agreement.cont",  align = "lll", digits = 2), include.rownames = T, size = "footnotesize")
@

@@

\begin{figure}
<<echo=FALSE>>=
plot(p.fixef.change)
@
\caption{Change in signficance of fixed effects meta-analysis pooled estimate after correction.}
\label{significance.change.fixed}
\end{figure}

>>>>

<<echo=FALSE>>=
library(knitr)
opts_chunk$set(fig.path='figure/ch02_fig',
               echo=TRUE, message=FALSE,
               fig.width=8, fig.height=3.7,
               out.width='\\textwidth-3cm',
               message=FALSE, fig.align='center',
               background="gray98", tidy=FALSE, #tidy.opts=list(width.cutoff=60),
               cache=TRUE
)
options(width=74)
@

@@

<<echo = FALSE, warning =FALSE>>=
#Merge comparisons with meta-analysis result:
mly.tomerge <- mly %>% select(meta.id, sig.Q, n.sig.type, n.sig.fishersz, sig.fixef, sig.ranef, sig.ranef.hkn)

data.mly <- merge(x = data.ext, y = mly.tomerge, by = c("meta.id")) #ideally, 278626 comparisons ( = sum(mly$n)), 
# but 300831 (18261 have smaller sample size than mly, but same meta.id)

data.mly <- data.mly %>% filter(total1 > 11 & total2 > 11) #Now 282550

sig.meta.c <- data.mly %>% select(sig.fixef, sig.ranef, sig.ranef.hkn) %>% 
  gather(key = "meta.analysis.method", value = "sig.meta.analysis") %>% mutate(sig.meta.analysis = factor(sig.meta.analysis))

sig.meta.count.c <- sig.meta.c %>% 
  group_by(sig.meta.analysis, meta.analysis.method) %>% count()

sig.meta.c <- sig.meta.c %>%  ggplot(aes(x = meta.analysis.method, fill = sig.meta.analysis)) + 
  geom_bar() + coord_flip() +
  theme_bw() + xlab(label = NULL) + scale_fill_discrete(labels= c("meta.nonsig", "meta.sig")) +
  annotate("text", x = sig.meta.count.c[sig.meta.count.c$sig.meta.analysis == 0,]$meta.analysis.method, y = 100000, 
           label = paste(sig.meta.count.c[sig.meta.count.c$sig.meta.analysis == 0,]$n, "nonsig."), 
           color = "white")

#Change in significance after meta-analysis, separated by significant effect size estimate
sig.meta <- data.mly %>% 
  select(meta.id, study.id, sig.fixef, sig.ranef, sig.ranef.hkn) %>% 
  gather(key = "meta.analysis.method", value = "sig.meta.analysis", sig.fixef:sig.ranef.hkn) %>% mutate(sig.meta.analysis = factor(sig.meta.analysis))

# duplicate.studies <- data.mly %>% group_by(file.nr, comparison.nr, outcome.nr, subgroup.nr, study.name) %>% 
#   count() %>% filter(n > 1) #Merging is not possible for these since studies bear multiple results, but are not distinguishable and are multiplied while merging.

sig.meta.plot <- merge(y = select(data.ext, meta.id, study.id, sig.type), 
                       x = sig.meta, by = c("meta.id", "study.id")) %>% 
  mutate(sig.type = factor(sig.type))

sig.meta.count <- sig.meta.plot %>% 
  group_by(sig.meta.analysis, meta.analysis.method, sig.type) %>% count()

sig.meta.plot1 <- sig.meta.plot %>% filter(sig.type == 1) %>%  ggplot(aes(x = meta.analysis.method, fill = sig.meta.analysis)) + 
  geom_bar() + coord_flip() + ggtitle("Significant studies") +
  theme_bw() + xlab(label = NULL) + scale_fill_discrete(labels= c("non-significant", "significant")) +
  annotate("text", x = sig.meta.count[which(sig.meta.count$sig.type == 1)[1:3],]$meta.analysis.method, y = 20000, 
           label = paste(sig.meta.count[which(sig.meta.count$sig.type == 1)[1:3],]$n, "changed to nonsig."), 
           color = "white")

nonsig.meta.plot <- sig.meta.plot %>% filter(sig.type == 0) %>%  ggplot(aes(x = meta.analysis.method, fill = sig.meta.analysis)) + 
  geom_bar() + coord_flip() + ggtitle("Non-significant studies") +
  theme_bw() + xlab(label = NULL) + scale_fill_discrete(labels= c("non-significant", "significant")) +
  annotate("text", x = sig.meta.count[which(sig.meta.count$sig.type == 0)[4:6],]$meta.analysis.method, y = 100000, 
           label = paste(sig.meta.count[which(sig.meta.count$sig.type == 0)[4:6],]$n, "changed to sig."), 
           color = "white")

#Change in significance after meta-analysis, separated by significant heterogeneity
sig.meta.Q <- data.mly %>% 
  select(meta.id, sig.Q, sig.fixef, sig.ranef, sig.ranef.hkn) %>% 
  gather(key = "meta.analysis.method", value = "sig.meta.analysis", sig.fixef:sig.ranef.hkn) %>% 
  mutate(sig.meta.analysis = factor(sig.meta.analysis))

sig.meta.count.Q <- sig.meta.Q %>% 
  group_by(sig.meta.analysis, meta.analysis.method, sig.Q) %>% count()

sig.meta.Q1 <- sig.meta.Q %>% filter(sig.Q == 1) %>%  ggplot(aes(x = meta.analysis.method, fill = sig.meta.analysis)) + 
  geom_bar() + coord_flip() + ggtitle("Significant heterogeneity Q studies") +
  theme_bw() + xlab(label = NULL) + scale_fill_discrete(labels= c("non-significant", "significant")) +
  annotate("text", x = sig.meta.count.Q[which(sig.meta.count.Q$sig.Q == 1)[1:3],]$meta.analysis.method, y = 20000, 
           label = paste(sig.meta.count.Q[which(sig.meta.count.Q$sig.Q == 1)[1:3],]$n, "nonsig."), 
           color = "white")

nonsig.meta.Q <- sig.meta.Q %>% filter(sig.Q == 0) %>%  ggplot(aes(x = meta.analysis.method, fill = sig.meta.analysis)) + 
  geom_bar() + coord_flip() + ggtitle("Non-significant heterogeneity Q studies") +
  theme_bw() + xlab(label = NULL) + scale_fill_discrete(labels= c("non-significant", "significant")) +
  annotate("text", x = sig.meta.count.Q[which(sig.meta.count.Q$sig.Q == 0)[1:3],]$meta.analysis.method, y = 100000, 
           label = paste(sig.meta.count.Q[which(sig.meta.count.Q$sig.Q == 0)[1:3],]$n, "nonsig."), 
           color = "white")

@

\section{Meta-analysis}
The data at hand is composed of over 400,000 comparisons with balanced sample size $n > 24$, which has been chosen as a threshold to assess the significance of treatment effect. The significance of the p-values of treatment effect estimates was calculated for all comparisons in the dataset with acceptable sample size and outcome measures as odds ratios, risk ratios, risk differences and mean and standardized mean differences. This led to somewhat more than 400,000 significance test results. Since most of the times, the study publication year is available, the fraction of significant treatment effects found is shown over time in Figure \ref{study.significance.overtime}.The $p$-value was chosen to be $0.05$  only times where a reasonbly large number of effect estimates is available is shown in order to reduce random fluctuation ($n > 800$).

\begin{figure}
<<echo=FALSE, warning=FALSE>>=
data.ext %>% filter(!is.na(sig.type)) %>%  ggplot(aes(x = study.year, fill = factor(sig.type), stat(count))) + geom_density(na.rm = T, position = "fill") + 
  labs(fill = "Significance") + scale_fill_discrete(labels= c("Yes", "No"))+ scale_x_continuous(limits = c(1970, 2017))
@
\caption{Mean of the absolute value of the normalized effect size plotted against the total sample size.}
\label{study.significance.overtime}
\end{figure}

It is possible to analyse all studies with one or more replica by meta-analysis. There are different meta-analys methods that can be applied. While fixed effects meta-analysis pools the estimates and their variance, random effects meta-analysis additionally estimates the between study variances and adds this variance to the pooled within-study variances when assessing the uncertainity of the pooled effect estimate. The Hartung and Knapp adjustment of p-values and confidence intervals for random effects meta-analysis is more conservative than random-effects meta-analysis and will yieild less significant results than the latter.

\vspace{0mm}
In the following, the study results are combined with the results of their corresponding meta-analysis. Thus, every study that has one or more replicate was analysed by meta-analysis. In the following, the significance of the study, the primary significance, is compared with the significance of the meta-analysis or secondary significance.

\vspace{0mm}
In Figure \ref{primary.secondary.significance}, the overall fraction of studies with significant and non-significant meta-analysis results is shown. The meta-analysis method used to pool the treatment effect estimates and assessing significance is indicated on the right hand side of the Figure. One can see the overall fraction of rejected null-hypotheses of no treatment effect.

\begin{figure}
<<echo=FALSE, warning=FALSE>>=
plot(sig.meta.c)
@
\caption{Overall fraction of studies whose treatment effect estimate was significant when pooled by means of meta-analysis. The fractions have been calculated
by fixed-effects, random-effects and Hartung and Knapp adjusted random-effects meta-analysis}
\label{primary.secondary.significance}
\end{figure}

One can visualise the fraction of studies with significant treatment effect separately for studies with significant primary treatment effect, i.e. non-pooled effect, and non-significant primary treatment effect. This can be seen in Figure \ref{primary.secondary.significance.sep.sig}. One can see that primary significance and non-significance is often overruled when performing a meta-analysis.

\begin{figure}
<<echo=FALSE, warning=FALSE>>=
grid.arrange(sig.meta.plot1, nonsig.meta.plot, ncol= 1)
@
\caption{Overall fraction of studies whose treatment effect estimate was significant when pooled by means of meta-analysis, separated by significance of study treatment effect estimate. The fractions have been calculated
by fixed-effects, random-effects and Hartung and Knapp adjusted random-effects meta-analysis}
\label{primary.secondary.significance}
\end{figure}

The separation of studies can also be made based on the significance of heterogeneity between them when pooling them by means of a meta-analysis. Significant heterogeneity between studies corresponds to a rejection of the null-hypothesis that all the study treatment effect estimates share the same underlying distribution. The test used to assess heterogeneity was based on the between-study heterogeneity estimate $Q$ estimated as \citet{tau.estimator} suggested. This is shown in Figure \ref{primary.secondary.significance.sep.sig}.

\begin{figure}
<<echo=FALSE, warning=FALSE>>=
grid.arrange(sig.meta.Q1, nonsig.meta.Q, ncol= 1)
@
\caption{Overall fraction of studies whose treatment effect estimate was significant when pooled by means of meta-analysis, separated by significance of study treatment effect estimate. The fractions have been calculated
by fixed-effects, random-effects and Hartung and Knapp adjusted random-effects meta-analysis}
\label{primary.secondary.significance.sep.sig}
\end{figure}

@@

\begin{figure}
<<echo=FALSE>>=
data %>% filter(total1 + total2 > 10 & total1 + total2 < 200) %>% 
  mutate(sample.size = total1 + total2, scaled.effect = abs(scale(effect, center = T, scale = T))) %>%
  group_by(sample.size) %>% 
  summarize(median.effect = median(scaled.effect, na.rm = T)) %>% 
  ggplot(aes(x = sample.size, y = median.effect)) + geom_point() + theme_bw() + xlab("sample size") + ylab("median absolute normalized effect size")
@
\caption{Median of the absolute value of the normalized effect size plotted against the total sample size.}
\label{effect.samplesize}
\end{figure}

@@

\begin{figure}
<<echo=FALSE>>=
data %>% filter(total1 + total2 > 10 & total1 + total2 < 100) %>% 
  filter(outcome.measure.new == "Risk Ratio" | outcome.measure.new == "Odds Ratio" | outcome.measure.new == "Mean Difference" |
           outcome.measure.new == "Std. Mean Difference") %>% 
  group_by(outcome.measure.new) %>% 
  mutate(sample.size = total1 + total2, scaled.effect = abs(scale(effect, center = T, scale = T))) %>%
  group_by(sample.size, outcome.measure.new) %>% 
  summarize(median.effect = median(scaled.effect, na.rm = T)) %>% 
  ggplot(aes(x = sample.size, y = median.effect, group = outcome.measure.new)) + geom_point() + facet_wrap(~outcome.measure.new) + 
  theme_bw() + xlab("sample size") + ylab("median absolute normalized effect size")
@
\caption{Median of the absolute value of the normalized effect size plotted against the total sample size, separated for outcome measures.}
\label{effect.samplesize.separated}
\end{figure}

A second way to analyse meta-analyses is a cumulative meta-analysis that should reveal shifts in treatment effect sizes over time. This can again be done for the entire dataset, i.e. for all effect estimates and their ordering in time. It is important to scale the effect estimates here with respect to the estimates of the replication studies that are about the same subject, have the same outcome measure, etc. Also, only effect estimates that can be compared to other estimates before are included, i.e. only study results with one or more replica are included. Time needs to be scaled and normalized in order to compare multiple time-trends in effect size to each other and gain insights in the overall trend. This is done in Figure \ref{effects.overtime.separated}, separared for odds rati, risk ratio, and mean and standardized difference outcome measures. In order to reduce the spread over time of effect sizes, only studies between 1970 and 2019 were included, as well as studies with a minimal group size of 12 participants (each group).

\begin{figure}
<<echo=FALSE>>=
data.ext %>% filter(outcome.measure.new == "Mean Difference" | outcome.measure.new == "Std. Mean Difference" | 
                outcome.measure.new == "Risk Ratio" | outcome.measure.new == "Odds Ratio") %>% 
  filter(total1 > 11 & total2 > 11) %>% filter(!is.na(study.year) & study.year > 1970 & study.year < 2019) %>% 
  group_by(meta.id) %>% mutate(n = n()) %>% filter(n > 1) %>% 
  mutate(scaled.effect = abs(scale(effect)), scaled.time = scale(study.year)) %>% 
  ggplot(aes(x = scaled.time, y = scaled.effect)) + geom_point(size = .5, alpha = 0.2)  + facet_wrap(~outcome.measure.new) +
  geom_smooth() + theme_bw() +  ylab("absolute scaled effect")  + ylab("scaled time")
@
\caption{Absolute values of scaled effect sizes over scaled time, separated for outcome measures.}
\label{effects.overtime.separated}
\end{figure}

% It is also possible to look how the amount of meta-analysis with significant pooled effect size estimates changes over time, e.g. mean publication year (of studies included in the meta-analysis). 0.05 is used here as the threshold for $p$-value significance and subsequent null-hypothesis rejection. To constrict random fluctuation in the proportion, only mean publication years with a certain amount of meta-analyses are shown ($n < 180$). The $p$-value of the fixed effects meta analysis model is used to decide about rejection of the null-hypothesis. As Figure \ref{treatment.effect.significance.overtime} shows, the proportion of meta-analyses with significant treatment effect estimates stays roughly constant over time. 
% 
% \begin{figure}
% <<echo=FALSE>>=
% meta %>% filter(mean.publication.year < 2013 & mean.publication.year > 1990) %>% 
%   ggplot(aes(x = mean.publication.year, fill = factor(sig.fixef), stat(count))) + geom_density(na.rm = T, position = "fill") + 
%   labs(fill = "Significance") + scale_fill_discrete(labels= c("Yes", "No"))
% @
% \caption{Proportion of significant treatment effect estimates over time (mean publication year).}
% \label{effect.samplesize}
% \end{figure}
%

@@

p.change.ranef.bin <- meta.bin %>% ungroup() %>% 
  select(sig.change.ranef.reg, sig.change.ranef.copas, sig.change.ranef.trimfill) %>% 
  gather(key = "correction.method", value = "sig.change") %>%  
  mutate(sig.change = factor(sig.change)) %>% 
  ggplot(aes(x = correction.method, fill = sig.change)) + geom_bar() + coord_flip() + ggtitle("Binary outcomes")+ 
  theme_bw() + xlab(label = NULL) + scale_fill_discrete(labels= c("unchanged.nonsig", "change.to.sig", "unchanged.sig", "change.to.nonsig")) +
  annotate("text", x = change.ranef.bin[change.ranef.bin$sig.change == 3,]$correction.method, y = 1000, 
         label = paste(change.ranef.bin[change.ranef.bin$sig.change == 3,]$n, "changed to nonsig."), 
         color = "white")

@@

p.change.ranef.cont <- meta.cont %>% ungroup() %>% 
  select(sig.change.ranef.reg, sig.change.ranef.copas, sig.change.ranef.trimfill) %>% 
  gather(key = "correction.method", value = "sig.change") %>%  
  mutate(sig.change = factor(sig.change)) %>% 
  ggplot(aes(x = correction.method, fill = sig.change)) + geom_bar() + coord_flip() + ggtitle("Continuous outcomes") +
  theme_bw() + xlab(label = NULL) + scale_fill_discrete(labels= c("unchanged.nonsig", "change.to.sig", "unchanged.sig", "change.to.nonsig")) +
  annotate("text", x = change.ranef.cont[change.ranef.cont$sig.change == 3,]$correction.method, y = 500, 
         label = paste(change.ranef.cont[change.ranef.cont$sig.change == 3,]$n, "changed to nonsig."), 
         color = "white")

@@

p.ranef.change <- meta %>% ungroup() %>% 
  select(sig.change.ranef.reg, sig.change.ranef.copas, sig.change.ranef.trimfill) %>% 
  gather(key = "correction.method", value = "sig.change") %>%  
  mutate(sig.change = factor(sig.change)) %>% 
  ggplot(aes(x = correction.method, fill = sig.change)) + geom_bar() + coord_flip() + 
  theme_bw() + xlab(label = NULL) + scale_fill_discrete(labels= c("unchanged.nonsig", "change.to.sig", "unchanged.sig", "change.to.nonsig")) +
  annotate("text", x = change.ranef[change.ranef$sig.change == 3,]$correction.method, y = 1500, 
         label = paste(change.ranef[change.ranef$sig.change == 3,]$n, "changed to nonsig."), 
         color = "white")

@@

p.change.fixef.bin <- meta.bin %>% ungroup() %>% 
  select(sig.change.fixef.reg, sig.change.fixef.copas, sig.change.fixef.trimfill) %>% 
  gather(key = "correction.method", value = "sig.change") %>%  
  mutate(sig.change = factor(sig.change)) %>% 
  ggplot(aes(x = correction.method, fill = sig.change)) + geom_bar() + coord_flip() + ggtitle("Binary outcomes")+  
  theme_bw() + xlab(label = NULL) + scale_fill_discrete(labels= c("unchanged.nonsig", "change.to.sig", "unchanged.sig", "change.to.nonsig")) +
  annotate("text", x = change.fixef.bin[change.fixef.bin$sig.change == 3,]$correction.method, y = 800, 
         label = paste(change.fixef.bin[change.fixef.bin$sig.change == 3,]$n, "changed to nonsig."), 
         color = "white")

@@

p.fixef.change <- meta %>% ungroup() %>% 
  select(sig.change.fixef.reg, sig.change.fixef.copas, sig.change.fixef.trimfill) %>% 
  gather(key = "correction.method", value = "sig.change") %>%  
  mutate(sig.change = factor(sig.change)) %>% 
  ggplot(aes(x = correction.method, fill = sig.change)) + geom_bar() + coord_flip() + 
  theme_bw() + xlab(label = NULL) + scale_fill_discrete(labels= c("unchanged.nonsig", "change.to.nonsig", "unchanged.sig", "change.to.sig")) +
  annotate("text", x = change.fixef[change.fixef$sig.change == 3,]$correction.method, y = 1500, 
         label = paste(change.fixef[change.fixef$sig.change == 3,]$n, "changed to nonsig."), 
         color = "white")

@@

There are tests that can be applied to find out if small study effects are present in the meta analysis. For the precise description, see the methods section. Application of the tests is only recommended if there are ten or more studies \citep{cochrane.handbook} that can be used, so all meta-analyses with less than ten studies have been excluded.

@@

\begin{figure}
<<echo=FALSE, warning=FALSE>>=
plot(p.cont)
@
\caption{Proportion where the null hypothesis of no small study effect is rejected based on the 5\% significance level for different tests (continuous outcomes).}
\label{bias.results.cont}
\end{figure}

\begin{figure}
<<echo=FALSE, warning=FALSE>>=
plot(p.bin)
@
\caption{Proportion where the null hypothesis of no small study effect is rejected based on the 5\% significance level for different tests (dichotomous outcomes).}
\label{bias.results.bin}
\end{figure}

@@

\begin{figure}
<<echo=FALSE, warning=FALSE>>=
grid.arrange(p3, p4, ncol = 2)
@
\caption{Proportion where the null hypothesis of no small study effect is rejected based on the 5\% significance level for different tests (continuous outcomes).}
\label{bias.results.cont.sep}
\end{figure}

\begin{figure}
<<echo=FALSE, warning=FALSE>>=
grid.arrange(p1, p2, ncol = 2)
@
\caption{Proportion where the null hypothesis of no small study effect is rejected based on the 5\% significance level for different tests (dichotomous outcomes).}
\label{bias.results.bin.sep}
\end{figure}

@@

\begin{figure}
<<echo=FALSE, warning=FALSE>>=
meta %>% filter(mean.publication.year < 2013 & mean.publication.year > 1990) %>% 
  ggplot(aes(x = mean.publication.year, fill = factor(thomson.test), stat(count))) + geom_density(na.rm = T, position = "fill") + 
  labs(fill = "Significance") + scale_fill_discrete(labels= c("Yes", "No"))
@
\caption{Proportion of test results where the null hypothesis of no small study effect is rejected over time (mean publication.year).}
\label{pub.bias.overtime}
\end{figure}

@@

<<results = 'asis', echo = FALSE, warning = FALSE>>=
print(xtable(x = t(binary.tests.agreement), caption = "Proportion of tests that agree in rejection ar acceptance of the null hypothesis that there is no small study effect (dichotomous outcomes)", label = "agreement.bin",  align = "lll", digits = 2), include.rownames = T, size = "footnotesize")
@

<<results = 'asis', echo = FALSE, warning = FALSE>>=
print(xtable(x = t(cont.tests.agreement), caption = "Proportion of tests that agree in rejection ar acceptance of the null hypothesis that there is no small study effect (continuous outcomes)", label = "agreement.cont",  align = "lll", digits = 2), include.rownames = T, size = "footnotesize")
@

@@

\begin{figure}
<<echo=FALSE, warning=FALSE>>=
plot(p.fixef.change)
@
\caption{Change in signficance of fixed effects meta-analysis pooled estimate after correction.}
\label{significance.change.fixed}
\end{figure}
